---
title: "CodeBook"
author: "Jerez"
date: "11/29/2020"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'asis')
```

### Course: Getting and Cleaning Data Course Project

One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained: [Recognition Using Smartphones](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones)

Here are the data for the project: [Project zip file](https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip)

We must create one R script called run_analysis.R that does the following.

1. Merges the training and the test sets to create one data set.
2. Extracts only the measurements on the mean and standard deviation for each measurement.
3. Uses descriptive activity names to name the activities in the data set
4. Appropriately labels the data set with descriptive variable names.
5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

## Library
```{r}
# Install and load required packages
if (!require('dplyr'))      {install.packages('dplyr')}
if (!require('data.table')) {install.packages('data.table')}

library(dplyr)
library(data.table)
`````

## Download the dataset

```{r}
# name for zip file
file.zip <- 'GCD_Final.zip' 

# Cheking if zip file exists
if (!file.exists(file.zip)){
 file.URL <- 'https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles...
 %2FUCI%20HAR%20Dataset.zip'
 download.file(file.URL, file.zip, method = 'curl')} 

# Checking if study folder exists
if (!file.exists("UCI HAR Dataset")) {unzip(file.zip)}
`````

# Reading the internal files

```{r}
# In the folder UCI HAR Dataset
act.lab   <- read.table("UCI HAR Dataset/activity_labels.txt", col.names = c("code", "activity"))
features  <- read.table("UCI HAR Dataset/features.txt",        col.names = c("n","functions"))

# In the sub-folder test of UCI HAR Dataset
subj.test <- read.table("UCI HAR Dataset/test/subject_test.txt", col.names = "subject")
x.test    <- read.table("UCI HAR Dataset/test/X_test.txt",       col.names = features$functions)
y.test    <- read.table("UCI HAR Dataset/test/y_test.txt",       col.names = "code")

# In the sub-folder train of UCI HAR Dataset
subj.train <- read.table("UCI HAR Dataset/train/subject_train.txt", col.names = "subject")
x.train    <- read.table("UCI HAR Dataset/train/X_train.txt",       col.names = features$functions)
y.train    <- read.table("UCI HAR Dataset/train/y_train.txt",       col.names = "code")
````

## 1. Merges the training and the test sets to create one data set.

```{r}
# We use the functions rbind & cbind

subj.data <- rbind(subj.train, subj.test)
x.data    <- rbind(x.train, x.test)
y.data    <- rbind(y.train, y.test)
merge.data<- cbind(subj.data, x.data, y.data)

#str(merge.data)
#head(merge.data)
````

## 2. Extracts only the measurements on the mean and standard deviation for each measurement.

```{r}
# We use select function and save the data like TidyData
# Also select the code of each subject (we use for the next part).

TidyData <- merge.data %>% select(subject, code, contains('mean'), contains('std'))
#head(TidyData)
````

## 3. Uses descriptive activity names to name the activities in the data set

```{r}
# The code of each subject is a number, we change for the activities who represent the number. 
TidyData$code[1:50] # example
print(act.lab[,2])

TidyData$code <- act.lab[TidyData$code, 2]
`````

## 4. Appropriately labels the data set with descriptive variable names.

For this part we need to read the variables of our TidyData and change for a descriptive name

Names of Feteatures will labelled using descriptive variable names.
 - Acc is replaced by Accelerometer
 - BodyBody is replaced by Body
 - Gyro is replaced by Gyroscope
 - Mag is replaced by Magnitude
 - gravity is replaced by Gravity
 - angle is replaced by Angle
 - prefix t is replaced by Time
 - prefix f is replaced by Frequency
 - tBody is replaced by TimeBody
 - -mean() is replaced by Mean
 - -std() is replaced by STD
 - -freq() is replaced by Frecuency

````{r}
names(TidyData)

names(TidyData)[2] = 'Activity'

names(TidyData) <- gsub("Acc", "Accelerometer", names(TidyData))
names(TidyData) <- gsub("BodyBody", "Body",     names(TidyData))
names(TidyData) <- gsub("Gyro", "Gyroscope",    names(TidyData))
names(TidyData) <- gsub("Mag", "Magnitude",     names(TidyData))
names(TidyData) <- gsub("gravity", "Gravity",   names(TidyData))
names(TidyData) <- gsub("angle", "Angle",       names(TidyData))
names(TidyData) <- gsub("^t", "Time",           names(TidyData))
names(TidyData) <- gsub("^f", "Frequency",      names(TidyData))
names(TidyData) <- gsub("tBody", "TimeBody",    names(TidyData))
names(TidyData) <- gsub("-mean()", "Mean",      names(TidyData), ignore.case = TRUE)
names(TidyData) <- gsub("-std()", "STD",        names(TidyData), ignore.case = TRUE)
names(TidyData) <- gsub("-freq()", "Frequency", names(TidyData), ignore.case = TRUE)

#head(TidyData)
names(TidyData)
````
## 5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

````{r}
Final.TidyData <- TidyData %>%
    group_by(subject, Activity) %>%
    summarise_all(funs(mean))

# Export the final data
write.table(Final.TidyData, "Final.TidyData.txt", row.name = FALSE)

# Cheking the data

str(Final.TidyData)
head(Final.TidyData)

`````

## Create run_analysis.R


````{r}
# knitr::purl('CodeBook.Rmd')

# By default its create a code whit the same name of CodeBook, in the folder you can change. 
````




